---
title: "Getting Started with Die Rolling"
author: "Shi Fan Jin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Die Rolling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
library(dieroller)
```

##Die Rolling
The package '"dieroller"' is a minimal implementation for simulating rolling a die a given number of times, and to visualize the relative frequencies in such a series of rolls.

##Creating a 'die'
The first step is to create a '"die"' object with the function 'die()':
```{r}
fair_die <- die()
fair_die
```
By default, 'die()' creates a fair die with standard 6 sides and with probability values for each side of 1/6.

You can change the defaults with arguments sides and prob:
```{r}
# create a loaded die
loaded_die <- die(prob = c(0.075, 0.1, 0.125, 0.15, 0.20, 0.35))
loaded_die

#die with non-standard sides
weird_die <- die(sides = c('i', 'ii', 'iii', 'iv', 'v', 'vi'))
weird_die
```
##Rolling a Die
Once you have defined a '"die"' object, you can use the function 'roll()' to roll a die a given number of 'times'. The output will be an object of class '"roll"', which will contain the vector of 'rolls':
```{r}
# roll fair die 50 times
fair50 <- roll(fair_die, times = 50)
fair50

# roll 50 times
set.seed(123)
fair_50rolls <- roll(fair_die, times = 50)
# print
fair_50rolls

# what's in fair50?
names(fair50)
fair50$rolls
fair50$sides
fair50$prob
fair50$total

# string die
str_die <- die(
  sides = c('a', 'b', 'c', 'd', 'e', 'f'),
  prob = c(0.075, 0.1, 0.125, 0.15, 0.20, 0.35))
# roll 20 times
set.seed(123)
str_rolls <- roll(str_die, times = 20)
names(str_rolls)
str_rolls
```
The 'summary()' function gives you a nice table of frequencies:
```{r}
set.seed(123)
fair50_sum <- summary(fair_50rolls)
fair50_sum

# what's in the summary
class(fair50_sum)
names(fair50_sum)
fair50_sum$freqs
```

##Plot of relative frequencies
You can use the function 'plot()' to create a graohic with the relative frequencies of a series of die rolls:
```{r}
plot(fair_50rolls)
```

##Additional Method
There are also some additional method of using the object 'die' and the function 'roll()', such as an extraction value for the given roll, a replacement method to replace a given value of a given roll, and addition method to add more rolls.
```{r}
# roll fair die
set.seed(123)
fair_die <- die()
fair500 <- roll(fair_die, times = 500)
# summary method
summary(fair500)

# extracting roll in position 500
fair500[500]

# replacing last roll
fair500[500] <- 1
fair500[500]
#checking
summary(fair500)

# adding 100 rolls
fair600 <- fair500 + 100
#checking
summary(fair600)

# plot method
plot(fair500, 500)
```
