---
title: "Hw02"
author: "Shi Fan Jin"
date: "3/4/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = '../images/')
```
# Please do not use code chunk options such as: echo = FALSE, eval = FALSE, results = 'hide'. All chunks must be visible and evaluated.

```{r}
library(dplyr)
library(ggplot2)
```

```{r}
shots_data <- read.csv(file = "../data/shots-data.csv", stringsAsFactors = FALSE)
```

#####5.1) Total Shots by Player (10 pts)
```{r}

total_shots <- arrange(
  summarise(
    group_by(shots_data, name),
    total = n()
  ),
  desc(total)
)

total_shots


```

#####5.2) Effective Shooting Percentage (10 pts)
```{r}

# Effective Shooting % by Player
shots_data %>% 
  group_by(name) %>%
  summarise(
    total = n(),
    made = sum(shot_made_flag == "made shot"),
    perc_made = made/total) %>%
  arrange(desc(perc_made))


# 2PT Effective Shooting % by Player
shots_data %>%
  group_by(name) %>%
  summarise( 
    total = n(),
    made = sum(shot_made_flag == "made shot" & shot_type == "2PT Field Goal"),
    perc_made = made/total) %>%
  arrange(desc(perc_made))
  

# 3PT Effective Shooting % by Player

shots_data %>%
  group_by(name) %>%
  summarise( 
    total = n(),
    made = sum(shot_made_flag == "made shot" & shot_type == "3PT Field Goal"),
    perc_made = made/total) %>%
  arrange(desc(perc_made))


```

#####6.1) dplyr table (10 pts)
```{r}
shot_distance <- shots_data %>%
  group_by(shot_distance) %>%
  summarise(
    total = n(),
    made = sum(shot_made_flag == "made shot"),
    made_shot_prop = made/total) %>%
  select(c(shot_distance, made_shot_prop))

shot_distance
```

#####6.2) ggplot (10 pts)
```{r}
ggplot(data = shot_distance, aes(x = shot_distance, y = made_shot_prop)) +
  geom_point() 

```
**We can see that for those shot_distance below 30 has a higher proportion, whereas for those shot from a longer distance, the chance it got in is very low which is around 0 in the graph. So, from my point of view, we can confirm that the shorter the distance, the more effective the shots. As for the distance of having a percentage (of making a shot) of 50% or more is will be the ones around 1 to 3, and also for 51(which in this case is a extreme value for me).**

#####7) Total number of shots by minute of occurrence (10 pts)
```{r}
shots_minute_occurrence <- shots_data %>%
  group_by(name, minute) %>%
  summarise(
    total = n())
 

ggplot(dat = shots_minute_occurrence, aes(x = minute, y = total)) +
  geom_point(col = "blue") +
  facet_wrap( ~ name) +
  theme_minimal() +
  geom_path(aes(x=minute,y=total),linejoin="round",lineend="round",col="blue", alpha = 0.3) +
  scale_x_continuous(breaks=c(1,12,24,36,48)) +
  geom_rect(xmin = 0, xmax = 12, ymin = 0, ymax = 60, fill = "grey80", alpha = 0.01) +
  geom_rect(xmin = 24, xmax = 36, ymin = 0, ymax = 60, fill = "grey80", alpha = 0.01) +
  ggtitle("Total number of shots (by minute of occurrence)") +
  ylab("total number of shots")

```

